---
title: "ISyE6414 Project"
author: "Team 14: Dataset for Future Prediction"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Read Data file in
```{r, message=FALSE, warning=FALSE}
# Import library you may need
library(car)
library(dplyr)
library(TSstudio)
library(ggplot2)
library(tibble)

# Read the data set
raw <- read.csv("Online Retail.csv")
head(raw)
```

#Data Discription
InvoiceNo: Invoice number. Nominal, a 6-digit integral number uniquely assigned to each transaction. If this code starts with letter 'c', it indicates a cancellation.

StockCode: Product (item) code. Nominal, a 5-digit integral number uniquely assigned to each distinct product.

Description: Product (item) name. Nominal.

Quantity: The quantities of each product (item) per transaction. Numeric.

InvoiceDate: Invice Date and time. Numeric, the day and time when each transaction was generated.

UnitPrice: Unit price. Numeric, Product price per unit in sterling.

CustomerID: Customer number. Nominal, a 5-digit integral number uniquely assigned to each customer.

Country: Country name. Nominal, the name of the country where each customer resides.
```{r}
colnames(raw, do.NULL = TRUE, prefix = "col")
num_row <- nrow(raw)
cat('Number of rows:',num_row)
```
#Convert nominal data to catagorical data
```{r}
raw$InvoiceNo<-as.factor(raw$InvoiceNo)
raw$StockCode<-as.factor(raw$StockCode)
raw$Description<-as.factor(raw$Description)
raw$CustomerID<-as.factor(raw$CustomerID)
raw$Country<-as.factor(raw$Country)
raw$InvoiceDate <- as.Date(raw$InvoiceDate, "%m/%d/%Y")
sapply(raw, class)
```
#Removing redundant rows of the data
```{r}
rdd_raw <- distinct(raw)
num_row_rdd <- nrow(rdd_raw)
removed <- num_row - num_row_rdd
cat('Number of redundant rows been removed:',removed)
```

#Removing columns that contain missing value(s)
```{r}
missing_raw <- na.omit(rdd_raw)
num_row_missing <- nrow(missing_raw)
removed2 <- num_row_rdd-num_row_missing
cat('Number of rows contains missing values been further removed:',removed2)
cleaned_df<-missing_raw
head(cleaned_df)
nrow(cleaned_df)
```
#Segment the Data into First (Initial) and Second (Final) 6 Months (i.e. May 31, 2011)
```{r}
max(which((cleaned_df$InvoiceDate=="2011-05-31")))
final_df<-cleaned_df[1:max(which((cleaned_df$InvoiceDate=="2011-05-31"))),]
initial_df<-cleaned_df[min(which((cleaned_df$InvoiceDate=="2011-06-01"))):nrow(cleaned_df),]
```

#Explore the data
##Count catagorical data

```{r}
count_unique<-rapply(cleaned_df[c(1,2,7,8)],function(x) length(unique(x)))
initial_unique<-rapply(initial_df[c(1,2,7,8)],function(x) length(unique(x)))
final_unique<-rapply(final_df[c(1,2,7,8)],function(x) length(unique(x)))
count_unique
initial_unique
final_unique
```
By counting the unique appearance of the data, we can find out that:
1. The data frame contains 22190 unique transactions.
2. The data frame contains 3684 kinds of unique products.
3. The data frame contains 4372 unique customers.
4. The data is collected across 37 countries.

##Time series plot of sales over entire time span
```{r}
all_date <- as.Date(cleaned_df$InvoiceDate)
quantity_over_time <- aggregate(cleaned_df$Quantity, by=list(all_date), sum)
qov_plot<-ts_plot(quantity_over_time,
        title = "Daily Sales Quantity",
        Xtitle = "Time",
        Ytitle = "Number of Items")
qov_plot
```
Plot aggregated daily sales quantity. We do not see abnormal behaviors across time.

##Order quantity from top 10 countries
```{r}
country_table<-table(cleaned_df$Country)
country_table<-sort(country_table)
ranked<-country_table[c(29:38)]
print(ranked)
barplot(ranked,main = "Order Quantity from Top 10 Countries",las=2,cex.names=0.6)
```
Ranking countries that has the most orders.
United Kingdom ordered the most, and they take a huge portion of the data. Might be careful while performing regression, since huge amount of orders came from UK.

##Number of sales and return for Initial Six Months
```{r}
initial_return<-nrow(initial_df[initial_df$Quantity<0,])
initial_sale<-nrow(initial_df[initial_df$Quantity>0,])
cat('the number of sales of first six months of  the transactions is ', initial_sale)
cat('\nthe number of returns of first six months of the transactions is ', initial_return)

```
##Transaction Frequency of First Six Months
```{r}
#Unique Customers in First Six Months
initial_customer_table<-table(initial_df$CustomerID)
initial_customer_all<-as.data.frame(initial_customer_table)
initial_customer=filter(initial_customer_all,initial_customer_all$Freq>0)

#First 6 Months of return data frame
initial_return_df<-initial_df[initial_df$Quantity<0,]
initial_return_table<-table(initial_return_df$CustomerID)
initial_customer_return<-as.data.frame(initial_return_table)
initial_return=filter(initial_customer_return,initial_customer_return$Freq>0)
#First 6 Months of sales data frame
initial_sale_df<-initial_df[initial_df$Quantity>0,]
initial_sale_table<-table(initial_sale_df$CustomerID)
initial_customer_sale<-as.data.frame(initial_sale_table)
initial_sale=filter(initial_customer_sale,initial_customer_sale$Freq>0)
#Histogram of First Six Months
ggplot(initial_customer, aes(x=Freq)) +
  geom_histogram(binwidth=10) +
  labs(title="Histogram of Transaction Frequency of First Six Months of Customers", y = "Density")
ggplot(initial_return, aes(x=Freq)) +
  geom_histogram(binwidth=10) +
  labs(title="Histogram of Returns Frequency of First Six Months of Customers", y = "Density")
ggplot(initial_sale, aes(x=Freq)) +
  geom_histogram(binwidth=10) +
  labs(title="Histogram of Sales Frequency of First Six Months of Customers", y = "Density")

```


#Stock Ordering Frequency of First 6 Months
```{r}
#Stockcode ordering df for First 6 Months
initial_SC_ordering_table<-table(initial_sale_df$StockCode)
initial_SC_ordering_df<-as.data.frame(initial_SC_ordering_table)
initial_SC_ordering=filter(initial_SC_ordering_df,initial_SC_ordering_df$Freq>0)
#Plot of Stock Ordering First 6 Months
ggplot(initial_SC_ordering, aes(x=Freq)) +
  geom_histogram() +
  labs(title="Histogram of Item Ordering Frequency of All Products for First Six Months", y = "Density")

```

#Dataframe creation for Customers in First 6 Months
```{r}
initial_df2<-initial_df%>%
  mutate(Income=Quantity*UnitPrice)
initial_sale_df2<-initial_sale_df%>%
  mutate(Income=Quantity*UnitPrice)
initial_return_df2<-initial_return_df%>%
  mutate(Income=Quantity*UnitPrice)

#Unique Customers in First 6 Months
df<-data.frame(unique(initial_df2$CustomerID))
colnames(df)<-'CustomerID'

##Features of Customer Overall Orders
#Total Number of Purchase Orders by Customer
OrderNo <- initial_sale_df%>%
  group_by(CustomerID,InvoiceNo)%>%
  summarise(OrderNo = n())
Orders_Unique <- OrderNo%>%
  group_by(CustomerID)%>%
  summarise(Orders_Unique = n())
df <- left_join(df,Orders_Unique,'CustomerID')

#Total Number of Return Orders by Customer
ReturnNo <- initial_return_df%>%
  group_by(CustomerID,InvoiceNo)%>%
  summarise(ReturnNo = n())
Returns_Unique <- ReturnNo%>%
  group_by(CustomerID)%>%
  summarise(Returns_Unique = n())
df <- left_join(df,Returns_Unique,'CustomerID')

#Total Number of Invoice's (Purchase and Return) by Customer
InvoiceNo <- initial_df%>%
  group_by(CustomerID,InvoiceNo)%>%
  summarise(InvoiceNo = n())
Invoice_Unique <- InvoiceNo%>%
  group_by(CustomerID)%>%
  summarise(Invoice_Unique = n())
df <- left_join(df,Invoice_Unique,'CustomerID')

##Total Item Features
#Total Items Purchased by Customer
Sale_Quantity <- initial_sale_df%>%
  group_by(CustomerID)%>%
  summarise(Total_Items_Purchased=sum(Quantity))
df <- left_join(df,Sale_Quantity,'CustomerID')

#Total Quantity Per Purchase Basket
df$Quantity_Basket=df$Total_Items_Purchased/df$Orders_Unique

#Total Items Returned by Customer
Return_Quantity <- initial_return_df%>%
  group_by(CustomerID)%>%
  summarise(Total_Items_Returned=sum(Quantity))
df <- left_join(df,Return_Quantity,'CustomerID')

#Total Quantity Per Return
#df$Quantity_Return=df$Total_Items_Returned/df$Returns_Unique

#Net Total Items by Customer
Net_Quantity <- initial_df%>%
  group_by(CustomerID)%>%
  summarise(Net_Total_Items=sum(Quantity))
df <- left_join(df,Net_Quantity,'CustomerID')

#Net Total Quantity Per Invoice
#df$Average_Net_Quantity = df$Net_Total_Items/df$Invoice_Unique

##Unique Item Features
#Total Number of Unique Items Purchased by Customer
Purchase <- initial_sale_df%>%
  group_by(StockCode,CustomerID)%>%
  summarise(Purchase = n())
Purchase_Unique <- Purchase%>%
  group_by(CustomerID)%>%
  summarise(Types_Items_Purchased=n())
df <- left_join(df,Purchase_Unique,'CustomerID')

#Total Number of Unique Items Per Purchase Basket
Unique_Item_Per_Basket <-OrderNo%>%
  group_by(CustomerID)%>%
  summarise(Unique_Item_Per_Basket = mean(OrderNo))
df <- left_join(df,Unique_Item_Per_Basket,'CustomerID')

#Total Number of Unique Items Returned by Customer
Cancellation <- initial_return_df%>%
  group_by(StockCode,CustomerID)%>%
  summarise(Cancellation = n())
Cancellation_Unique <- Cancellation%>%
  group_by(CustomerID)%>%
  summarise(Types_Items_Returned=n())
df <- left_join(df,Cancellation_Unique,'CustomerID')

#Total Number of Unique Items Per Return Order
Unique_Item_Per_Return <-ReturnNo%>%
  group_by(CustomerID)%>%
  summarise(Unique_Item_Per_Return = mean(ReturnNo))
df <- left_join(df,Unique_Item_Per_Return,'CustomerID')

##Total Revenue Features
#Total Sales Revenue per Customer
Sales_Revenue <- initial_sale_df2%>%
  group_by(CustomerID)%>%
  summarise(Sales_Revenue=sum(Income))
df <- left_join(df,Sales_Revenue,'CustomerID')

#Average Order Value
#df$Average_Order_Value = df$Sales_Revenue/df$Orders_Unique

#Total Return Refund per Customer
Return_Refund <- initial_return_df2%>%
  group_by(CustomerID)%>%
  summarise(Return_Refund=sum(Income))
df <- left_join(df,Return_Refund,'CustomerID')

#Average Return Value
#df$Average_Return_Value = df$Return_Refund/df$Returns_Unique

#Total Revenue (Sales-Refund) per Customer
df$Total_Revenue=df$Sales_Revenue+df$Return_Refund

#Average Invoice Value
#df$Average_Invoice_Value = df$Total_Revenue/df$Invoice_Unique

##Features for per unit pricing
#Average Unit Price per Item Purchased
Average_Unit_Price_Purchase <- initial_sale_df2%>%
  group_by(CustomerID)%>%
  summarise(Average_Unit_Price_Purchase=sum(Income)/sum(Quantity))
df <- left_join(df,Average_Unit_Price_Purchase,'CustomerID')

#Average Unit Refund per Item Returned
Average_Unit_Refund_Return <- initial_return_df2%>%
  group_by(CustomerID)%>%
  summarise(Average_Unit_Refund_Return=sum(Income)/sum(Quantity))
df <- left_join(df,Average_Unit_Refund_Return,'CustomerID')

#Average Unit Revenue per Item in Invoice
#df$Average_Unit_Revenue=df$Total_Revenue/df$Net_Total_Items 

#Add Country Designation
Country <- initial_df%>%
  group_by(CustomerID)%>%
  summarise(Country=unique(Country))
df <- left_join(df,Country,'CustomerID')
```
```{r}
##Add Most popular item by Customer
#Most_Purchased_StockCode <- initial_sale_df2%>%
#  group_by(CustomerID,StockCode)%>%
#  summarise(Sum_Return=sum(Income))
#colnames(Most_Purchased_StockCode)[which(colnames(Most_Purchased_StockCode) %in% c("StockCode") )] <- #c("Most_Purchased_StockCode")
#Most_Purchased_StockCode <- Most_Purchased_StockCode %>% 
#  group_by(CustomerID) %>% filter(Sum_Return == max(Sum_Return))
#df <- left_join(df,Most_Purchased_StockCode[c('CustomerID','Most_Purchased_StockCode')],'CustomerID')
```
```{r}
#Is a customer purchasing the most profitable items
Profitable_items <- initial_sale_df2%>%
  group_by(StockCode)%>%
  summarise(Total_Monetary_Value=sum(Income))
Profitable_items<-Profitable_items[order(-Profitable_items$Total_Monetary_Value),]
# Top 3 items' StockCode: "23166", "22423", "85123A"
Is_Most_Popular <- initial_sale_df2 %>%
  mutate(Is_Most_Popular  = case_when(
    StockCode == "23166" ~ 1,
    StockCode == "22423" ~ 1,
    StockCode == "85123A" ~ 1,
    TRUE ~ 0
    ))
Is_Buying_Most_Popular <- Is_Most_Popular%>%
  group_by(CustomerID)%>%
  summarise(Is_Buying_Most_Popular=sum(Is_Most_Popular))
  
Is_Buying_Most_Popular <- Is_Buying_Most_Popular %>%
  mutate(Is_Buying_Most_Popular = 
           case_when(
             Is_Buying_Most_Popular>=1~1,
             Is_Buying_Most_Popular<1~0))

df <- left_join(df,Is_Buying_Most_Popular,'CustomerID')

df[is.na(df)] = 0

```


##After creating raw data set on customer level, we will start create additional predicting variables. Since we are studying RFM, we need to first calculate for Recency, Frequency, and Monetory value score. 
# https://towardsdatascience.com/recency-frequency-monetary-model-with-python-and-how-sephora-uses-it-to-optimize-their-google-d6a0707c5f17
# quantile method is borrowed from the link above.
###Recency is defined as: quantile of number of month since last purchase within all customers
```{r}
#Recency calculation for each customer. Number of days since last purchase to the end date of the initial data file (31 May 11)
Recency <- initial_df%>%
  group_by(CustomerID)%>%
  summarise(Recency=max(initial_df$InvoiceDate)-max(InvoiceDate))
Recency_Quantile <- Recency %>% mutate(Recency_Quantile = ntile(Recency, 100)/100)
df <- left_join(df,Recency_Quantile,'CustomerID')
```
###Frequency is originally defined as: quantile of number of purchases within the first 6 months within all customers
```{r}
df <- df %>% mutate(Frequency_Quantile = ntile(Invoice_Unique, 100)/100)
colnames(df)[which(colnames(df) %in% c("Invoice_Unique") )] <- c("Frequency")
```
###Monetory value is defined as: the highest value of all purchases by the customer expressed as a multiple of some benchmark value.
```{r}
df <- df %>% mutate(Monetory_Value_Quantile = ntile(Total_Revenue, 100)/100)
colnames(df)[which(colnames(df) %in% c("Total_Revenue") )] <- c("Monetory_Value")
```
### RFM score
```{r}
df$RFM_Score <- df$Recency_Quantile+df$Frequency_Quantile+df$Monetory_Value_Quantile
df <- df %>% mutate(RFM_Score = ntile(RFM_Score, 3))

```



##Dataframe for Last 6 Months Customer Revenue
```{r}
#Response for Revenue of Customers
temp <- final_df%>%
  mutate(Income=Quantity*UnitPrice)
Y_Income <- temp %>%
  group_by(CustomerID)%>%
  summarise(Y_Income = sum(Income))

```
##Combined Dataframe for Model
```{r}
final_data <- df
#final_data <- na.omit(final_data)
#final_data[, c(1: ncol(final_data))] <- sapply(final_data[, c(1: ncol(final_data))], as.numeric)
#final_data <- final_data[!is.infinite(rowSums(final_data)),]
final_data$Country<-as.factor(final_data$Country)
final_data$Is_Buying_Most_Popular<-as.factor(final_data$Is_Buying_Most_Popular)

final_data[ ,c('CustomerID','Invoice_Unique','Net_Total_Items','RFM_Score')] <- list(NULL)

write.csv(final_data,"2nd_months_predicting_variables.csv",row.names = FALSE)
```
```